Journey: A Study of Scalability of Memory Management in Linux

### 2017-05-05 paper.tex
a few more changes

\title{Journey: A Study of Scalabilty of Memory Management in Linux}
\author{Mark Mansi, Suhas Pai, Hasnain Ali Pirzada\\\texttt{\{markm, spai, hp\}@cs.wisc.edu}}
\date{}

@@

A similar problem was encountered by researchers studying both large storage
drives and large distributed storage systems. David is a system that allows
storage and big data researchers to run benchmarks requiring terabytes of
storage using much smaller storage devices.
David physically stores only
metadata and discards the contents of files. Reads to the disk causes David to
generate data on the fly. This design is based on the observation that
most benchmarking frameworks do not care about the actual content of the files,
and that most of the storage capacity of a drive tends to be data rather than
metadata \cite{david}. The Exalt system uses a similar methodology for
large-scale distributed storage systems \cite{exalt}. This methodology provides
a promising direction for large-memory system studies. One may consider ignoring
the contents of a process's heap and only storing kernel data structures and
process code and stack segments. However, the common use of custom data
structures makes generating heap contents on the fly
difficult. Moreover, on Linux, the kernel data structures for 1TB
of memory may also exceed the size of physical memory on current systems
\cite{simics}.

A virtualized environment can be used to provide a guest system with more memory
than is physically available to the host. A study of the limits of the KVM
hypervisor found that there is no fundamental limit to the size of guest
physical memory other than the hardware address width. However, currently, a
Linux host running KVM will require the guest memory to be backed by host
physical memory or host swap space \cite{ibmkvm}.  This means that when the
virtual machine uses the whole amount of memory allocated to it, the host can
swap pages to disk, resulting in poor performance, which can cause inaccurate
performance measurements and make large
benchmarks impractical.

@@

While priliminary, we believe that our study is a first step towards building tools 
to study large memory systems in a principled and informed manner.

>>>>

\title{Journey: A Study of Scalability of Memory Management in Linux}
\author{Mark Mansi, Suhas Pai, Hasnain Ali Pirzada\\\texttt{\{markm, spai, hp\}@cs.wisc.edu}}
\date{}

@@

A similar problem was encountered by researchers studying both large storage
drives and large distributed storage systems. David is a system that allows
storage and big data researchers to run benchmarks requiring terabytes of
storage using much smaller storage devices.
David physically stores only
metadata and discards the contents of files. Reads to the disk causes David to
generate data on the fly. This design is based on the observation that
most benchmarking frameworks do not care about the actual content of the files,
and that most of the storage capacity of a drive tends to be data rather than
metadata \cite{david}. The Exalt system uses a similar methodology for
large-scale distributed storage systems \cite{exalt}. This methodology provides
a promising direction for large-memory system studies. One may consider ignoring
the contents of a process's heap and only storing kernel data structures and
process code and stack segments. However, the common use of custom data
structures makes generating heap contents on the fly
difficult. Moreover, on Linux, the kernel data structures for 1TB
of memory may also exceed the size of physical memory on current systems
\cite{simics}. As our results show, in some cases, the overheads of
\texttt{vm\_area\_struct} alone could use significant amounts of memory.

A virtualized environment can be used to provide a guest system with more memory
than is physically available to the host. A study of the limits of the KVM
hypervisor found that there is no fundamental limit to the size of guest
physical memory other than the hardware address width. However, currently, a
Linux host running KVM will require the guest memory to be backed by host
physical memory or host swap space \cite{ibmkvm}.  This means that when the
virtual machine uses the whole amount of memory allocated to it, the host can
swap pages to disk, resulting in poor performance, which can cause inaccurate
performance measurements and make large
benchmarks impractical. Our findings indicate several areas of improvement that
are needed before such an approach could be practical.

@@

While priliminary, we believe that our study is a first step towards building
tools to study large memory systems in a principled and informed manner. As our
results show, approaches similar to David \cite{david, exalt} will need to deal
with high kernel data structure overheads. Meanwhile, virtualization-based tools
would need to deal with the poor scalability of swapping.
